---
title: "1. Data formatting"
format: html
---

## 1.1. Contaminants list and properties formatting

```{r}
rstudioapi::navigateToFile("../data-raw/1.1.data_formatting_contaminants.R")
```

Outputs :

-   data/contam_prop
-   data/PCB_ALL
-   data/PCB_ALL_lab
-   data/log_Kow_PCB_ALL
-   data/PCB_indicators
-   data/PCB_dioxineL
-   data/inter_dioxineL
-   data/HBCDD_ALL
-   data/HBCDD_ALL_lab
-   data/log_Kow_HBCDD_ALL
-   data/PFAS_ALL
-   data/PFAS_ALL_lab
-   data/subfamilies_PFAS_ALL
-   data/n_C_PFAS_ALL

## 1.2. Benthos data formatting

```{r}
rstudioapi::navigateToFile("../data-raw/1.2.data_formatting_benthos.R")
```

Outputs :

- data-raw/benthos_contam.csv
- data/benthos_metadata
- data/benthos_contam

## 1.3. Soles data formatting

```{r}
rstudioapi::navigateToFile("../data-raw/1.3.data_formatting_soles.R")
```

Outputs :

- data-raw/soles_contam.csv
- data/soles_contam


## 1.5. Compound selection

```{r}
rstudioapi::navigateToFile("../data-raw/1.4.compound_selection.R")
```

Outputs :

- data-raw/detection_rates.csv
- data/PCB
- data/PFAS
- data/HBCDD

## 1.6. Contaminant concentrations for selected compounds

```{r}
rstudioapi::navigateToFile("../data-raw/1.5.data_computation_selected_compounds.R")
```

Outputs :

- data-raw/benthos_contam.csv
- data-raw/soles_contam.csv
- data/benthos_contam
- data/soles_contam

- data/PCB_lab
- data/log_Kow_PCB

- data/PFAS_lab
- data/subfamilies_PFAS
- data/n_C_PFAS

- data/HBCDD_lab
- data/log_Kow_HBCDD

## 1.7. Stomach content

```{r}
rstudioapi::navigateToFile("../data-raw/1.6.data_stomach_content.R")
```

Outputs :

- data-raw/stomac_soles.csv
- data/stomac_soles
- data/table_species

## 1.8. Isotopic data

```{r}
rstudioapi::navigateToFile("../data-raw/1.7.data_isotopes.R")
```

Outputs :

- data-raw/isotopes.csv
- data/isotopes
