---
title: "2.Data_analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2.Data_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
format: 
  html:
    self-contained: true
---

```{r}
#| message: false
#| include: false
library(chopin.magnification)
library(tidyverse)
library(writexl)
library(patchwork)
library(ggrepel)
library(ggbreak)
library(FSA)
```

## 1. Diet composition

### Data computation:
```{r}
# rstudioapi::navigateToFile("../data-raw/2.1.diet_composition.R")
```

### Pie chart:
```{r}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
colors = gg_color_hue(4)

diet <- diet %>%
  arrange(desc(taxon)) %>%
  mutate(
    ypos = diet_portion / 2 + c(0, cumsum(diet_portion)[-length(diet_portion)]),
    label = paste0(round(diet_portion * 100), "%")
  )

ggplot_diet <- ggplot(data = diet) +
  aes(x = "", y = diet_portion, fill = taxon) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0) +
  scale_fill_manual(values = c(colors[4], colors[3], colors[2])) +
  labs(fill = "Taxon") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  ) +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = ypos, label = label), size = 5)

# Save plot
ggsave(plot = ggplot_diet, filename = "../inst/results/Data_analysis/1.ggplot_diet.jpg",
       width = 5, height = 3)
```

## 2. Lipid contents

Dataset:
```{r}
lipid <- contam |>
  select(species, lipid_percent_dw) |>
  full_join(metadata) |> 
  drop_na()
```

Descriptive statistics:
```{r}
min(lipid$lipid_percent_dw) # 0.75
max(lipid$lipid_percent_dw) # 7
lipid |> summarise(median(lipid_percent_dw), .by = grp)
```


Tests between taxa:
```{r}
dunn_lipid <- dunnTest(lipid_percent_dw ~ grp, data = lipid, method = "bonferroni")
res_dunn_grp <- dunn_lipid$res |> select(Comparison, P.adj) |> 
  mutate(P.adj = round(P.adj, digits = 3))
res_dunn_grp
```


```{r}
# Annotations for boxplots
counts <- lipid |>
  group_by(grp) |> 
  summarise(n = sum(!is.na(lipid_percent_dw)))

# Boxplots
ggplot_lipid <- ggplot(lipid) +
  aes(x = grp, y = lipid_percent_dw, fill = grp) +
  geom_boxplot() +
  labs(x = NULL, y = "Lipid content (%dw)", fill = "Taxa") +
  geom_text(
    data = counts,
    aes(x = c(1, 2, 3, 4),
        y = max(lipid$lipid_percent_dw)*1.1,
        label = paste0("n=", n)),
    inherit.aes = FALSE, hjust = 0.5, vjust = 1
  )

# Save plot
ggsave(plot = ggplot_lipid, filename = "../inst/results/Data_analysis/2.ggplot_lipid.jpg",
       width = 5, height = 3)
```


## 3. Total levels by family

Table of statistics for each taxon and feeding mode in ng/g dw, ng/g ww and ng/d lw:
```{r}
sub_data <- list("Soles" = contam |> filter(species == "Solea_solea"), 
                 "Prey" = contam |> filter(species != "Solea_solea"), 
                 "Bivalves" = contam |> filter(grp == "Bivalvia"), 
                 "Crustaceans" = contam |> filter(grp == "Crustacea"), 
                 "Polychaetes" = contam |> filter(grp == "Polychaeta"), 
                 "Suspensivores" = contam |> filter(alim == "Suspensivore"), 
                 "Deposivores" = contam |> filter(alim == "Deposivore"), 
                 "Omnivores" = contam |> filter(alim == "Omnivore"))
```


```{r}
families <- c("PCB", "PFAS")

sum_levels_ng_gdw <- c()
for(family in families){
  for(type in 1:length(sub_data)) {
  sum_levels_ng_gdw <- rbind(sum_levels_ng_gdw,
                             get_stats(type = names(sub_data)[type],
                                       contam = paste0(family, " (ng/gdw)"),
                                       sub_data[[type]][, paste0("somme", family, "_ng_gdw")] |> pull()))
  }
}

sum_levels_ng_gww <- c()
for(family in families){
  for(type in 1:length(sub_data)) {
  sum_levels_ng_gww <- rbind(sum_levels_ng_gww,
                             get_stats(type = names(sub_data)[type],
                                       contam = paste0(family, " (ng/gww)"),
                                       sub_data[[type]][, paste0("somme", family, "_ng_gww")] |> pull()))
  }
}

sum_levels_ng_glw <- c()
for(family in families[c(1)]){
  for(type in 1:length(sub_data)) {
  sum_levels_ng_glw <- rbind(sum_levels_ng_glw,
                             get_stats(type = names(sub_data)[type],
                                       contam = paste0(family, " (ng/glw)"),
                                       sub_data[[type]][, paste0("somme", family, "_ng_glw")] |> pull()))
  }
}

# List of data per unit
sum_levels <- list("sum_levels_ng_gdw" = sum_levels_ng_gdw, 
                   "sum_levels_ng_gww" = sum_levels_ng_gww, 
                   "sum_levels_ng_glw" = sum_levels_ng_glw)

# Save dataset
write_xlsx(x = sum_levels, path = "../inst/results/Data_analysis/3.sum_levels.xlsx")
```

## 4. Total levels by taxon and feeding mode

Plot of the total levels (sum in ng/g dw) in the different taxa and feeding mode:
```{r}
ggdata_sumfamily_taxalim <- contam |> 
  select(species, grp, alim, sommePCB_ng_gdw, sommePFAS_ng_gdw,) |> 
  pivot_longer(cols = c(sommePCB_ng_gdw, sommePFAS_ng_gdw), 
                names_to = "family", values_to = "sum_ng_gdw") |> 
  mutate(family = family |>  str_remove("somme"),
         family = family |>  str_remove("_ng_gdw"),
         family = factor(family, levels = c("PCB", "PFAS")))

# Annotations for plot
counts <- ggdata_sumfamily_taxalim |>
  group_by(family, grp) |> 
  summarise(counts = sum(!is.na(sum_ng_gdw)))
maxs <- ggdata_sumfamily_taxalim |>
  group_by(family) |> 
  summarise(maxs = max(sum_ng_gdw, na.rm = TRUE))
gg_text <- full_join(counts, maxs)

ggplot_sumfamily_taxalim <- ggplot(ggdata_sumfamily_taxalim) +
  aes(x = grp, y = sum_ng_gdw, fill = alim, col = alim) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.7) +
  facet_wrap(vars(family), nrow = 1, scales = "free_y") +
  # add median value
  stat_summary(inherit.aes = FALSE, aes(x = grp, y = sum_ng_gdw),
    fun = median, geom = "crossbar", size = 0.5, width = .5) +
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1)) +
  theme(legend.position = "bottom") +
  labs(x = NULL, y = "Sum concentrations (ng/g dw)", fill = "Feeding mode") +
  guides(color = FALSE) +
  geom_text(
    data = gg_text,
    aes(x = rep.int(1:4, times = 2),
        y = maxs*1.1,
        label = paste0("n=", counts)),
    inherit.aes = FALSE, hjust = 0.5, vjust = 1
  )

ggsave(
  filename = "../inst/results/Data_analysis/4.ggplot_contamination_sumfamily_taxalim.jpg",
  plot = ggplot_sumfamily_taxalim, width = 20, height = 10, units = "cm"
)
```

Test taxa effect:

```{r}
# PCB in ng/g lw
kruskal.test(sommePCB_ALL_ng_glw ~ grp, data = contam) # p-value < 0.05
# PFAS in ng/g dw
kruskal.test(sommePFAS_ALL_ng_gdw ~ grp, data = contam) # p-value < 0.05
```

```{r}
dunn_PCB_grp <- dunnTest(sommePCB_ALL_ng_glw ~ grp, data = contam |> drop_na(sommePCB_ALL_ng_glw))
dunn_PFAS_grp <- dunnTest(sommePFAS_ALL_ng_gdw ~ grp, data = contam |> drop_na(sommePFAS_ALL_ng_gdw))

res_dunn_grp <- dunn_PCB_grp$res |> select(Comparison, P.adj) |>
  full_join(dunn_PFAS_grp$res  |> select(Comparison, P.adj), by = "Comparison") |> 
  mutate(across(.cols = c(P.adj.x, P.adj.y), .fns = ~ signif(.x, digits = 4)))
names(res_dunn_grp) <- c("Comparison", "PCB", "PFAS")

writexl::write_xlsx(x = res_dunn_grp, path = "../inst/results/Data_analysis/4.test_dunn_taxa.xlsx")
```

Test feeding mode effect:

```{r}
# PCB in ng/g lw
kruskal.test(sommePCB_ALL_ng_glw ~ alim, data = contam) # p-value = 0.14
# PFAS in ng/g dw
kruskal.test(sommePFAS_ALL_ng_gdw ~ alim, data = contam) # p-value < 0.05
```

```{r}
dunn_PCB_alim <- dunnTest(sommePCB_ALL_ng_glw ~ alim, data = contam |> drop_na(sommePCB_ALL_ng_glw))
dunn_PFAS_alim <- dunnTest(sommePFAS_ALL_ng_gdw ~ alim, data = contam |> drop_na(sommePFAS_ALL_ng_gdw))

res_dunn_alim <- dunn_PCB_alim$res |> select(Comparison, P.adj) |>
  full_join(dunn_PFAS_alim$res  |> select(Comparison, P.adj), by = "Comparison") |> 
  mutate(across(.cols = c(P.adj.x, P.adj.y), .fns = ~ signif(.x, digits = 4)))
names(res_dunn_alim) <- c("Comparison", "PCB", "PFAS")

writexl::write_xlsx(x = res_dunn_alim, path = "../inst/results/Data_analysis/4.test_dunn_alim.xlsx")
```

# 5. Total levels by species

Plot of the total levels (sum in ng/g dw) in the different species:
```{r}
ggdata_sumfamily_species <- contam |> 
  select(labels, grp, alim, sommePCB_ng_gdw, sommePFAS_ng_gdw) |> 
  pivot_longer(cols = c(sommePCB_ng_gdw, sommePFAS_ng_gdw), 
                names_to = "family", values_to = "sum_ng_gdw") |> 
  mutate(family = family |>  str_remove("somme"),
         family = family |>  str_remove("_ng_gdw"),
         family = factor(family, levels = c("PCB", "PFAS")))

ggplot_sumfamily_species <- ggplot(ggdata_sumfamily_species) +
  aes(x = labels, y = sum_ng_gdw, fill = grp, color = grp, shape = alim) +
  geom_point() +
  facet_wrap(vars(family), nrow = 1, scales = "free_y") +
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1)) +
  theme(legend.position = "bottom") +
  labs(x = NULL, y = "Sum concentrations (ng/g dw)", color = "Taxon", shape = "Feeding mode") +
  guides(fill = FALSE, shape = guide_legend(nrow=2), color = guide_legend(nrow=2))


ggsave(
  filename = "../inst/results/Data_analysis/5.ggplot_contamination_sumfamily_species.jpg",
  plot = ggplot_sumfamily_species, width = 20, height = 10, units = "cm"
)
```

# 6. Total levels in PCBs versus lipid content

There is a significant positive correlation between lipid content and PCB contamination levels (Spearman's test: p-value = 5×10⁻⁴, ρ = 0.43). However, some species appear to deviate from this general trend. For example, Crangon crangon shows highly variable lipid levels without a corresponding increase in PCB contamination. Conversely, Limecola balthica exhibits relatively stable lipid content despite a wide range of PCB concentrations.

```{r}
cor.test(x = contam$lipid_percent_dw, y = contam$sommePCB_ng_gdw, method = "spearman")
```

```{r}
ggplot_lipid_PCB <- ggplot(contam) +
  aes(x = lipid_percent_dw, y = sommePCB_ng_gdw, color = labels) +
  geom_point() +
  scale_color_manual(values = my_colors) +
  labs(title = "PCB", x = "Lipid content (%dw)", y = "Sum concentrations (ng/g dw)", color = NULL)
ggplot_lipid_PCB
```

```{r}
ggsave("../inst/results/Data_analysis/6.lipid_vs_contaminants.png", ggplot_lipid_PCB, width = 15, height = 12, unit = "cm")
```

# 7. Average contamination profiles in soles and their prey

```{r}
contam_profile <- contam |> 
  mutate(type = case_when(species == "Solea_solea" ~ "Sole",
                          TRUE ~ "Prey")) |> 
  select(species, type, ends_with("normalised_sum_ng_gdw")) |> 
  group_by(type) |> 
  summarise(across(.cols = -c(species), .fns = ~ mean(.x, na.rm = TRUE)*100), .groups = "drop") |> 
  pivot_longer(cols = -type, names_to = "chemical", values_to = "percent") |> 
  mutate(chemical = chemical |> str_remove("_normalised_sum_ng_gdw"),
         family = case_when(str_starts(chemical, "CB") ~ "PCB",
                            TRUE ~ "PFAS")) |> 
  mutate(family = factor(family, levels = c("PCB", "PFAS")),
         type = factor(type, levels = c("Sole", "Prey"))) |> 
  group_by(family)  |> 
  mutate(chemical = fct_reorder(chemical, percent, .fun = min)) |>  # order chemical by percent values
  ungroup()

ggplot_profile <- ggplot(contam_profile) +
  aes(x = chemical, y = percent, fill = type) +
  geom_col(position = "dodge") +
  facet_wrap(vars(family), nrow = 1, scales = "free_y") +
  labs(x = NULL, y = "Percentage of total concentrations (ng/g dw)", fill = NULL) +
  theme(legend.position = "bottom") +
  coord_flip()
ggplot_profile

ggsave("../inst/results/Data_analysis/7.contamination_profiles.jpg", ggplot_profile, width = 8, height = 4, dpi = 300)
```

```{r}
contam_profile_report <- contam_profile |> 
  mutate(percent = round(percent, digits = 1)) |> 
  pivot_wider(names_from = type, values_from = percent) |> 
  select(-family)

writexl::write_xlsx(x = contam_profile_report, path = "../inst/results/Data_analysis/7.contamination_profile.xlsx")
```

# 8. Physico-chemical properties and contamination profiles

```{r}
contam_profile_prop <- left_join(contam_profile, contam_prop)
```

```{r}
contam_profile_prop_PCB <- contam_profile_prop |> 
  filter(family == "PCB", type == "Sole") |>
  mutate(sub_family_TAG = factor(sub_family_TAG, 
                                 levels = c("tri-CB", "tetra-CB", "penta-CB", 
                                            "hexa-CB", "hepta-CB")))

ggplot_contam_profile_prop_PCB <- ggplot(contam_profile_prop_PCB) +
  aes(x = logKow, y = percent, size = sub_family_TAG, color = sub_family_TAG) +
  geom_point() +
  geom_text_repel(aes(label = chemical), size = 3, 
                  point.padding = 1, max.overlaps = 20) +
  theme(legend.position = "bottom") +
  theme(
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8)
  ) +
  labs(title = "PCB", y = "Percentage of total concentrations (ng/g dw)", 
       size = "Chlorine substitution", color = "Chlorine substitution")
ggplot_contam_profile_prop_PCB

ggsave("../inst/results/Data_analysis/8.contam_profiles_prop_PCB.jpg",
       ggplot_contam_profile_prop_PCB, width = 6, height = 5, dpi = 300)
```


```{r}
contam_profile_prop_PFAS <- contam_profile_prop |> 
  filter(family == "PFAS", type == "Sole") |>
  mutate(n_C = as.numeric(n_C))

ggplot_contam_profile_prop_PFAS <- ggplot(contam_profile_prop_PFAS) +
  aes(x = n_C, y = percent, size = sub_family_TAG, color = sub_family_TAG) +
  geom_point() +
  geom_text_repel(aes(label = chemical), size = 3,
                  point.padding = 1, max.overlaps = 20) +
  labs(title = "PFAS", y = "Percentage of total concentrations (ng/g dw)",
       x = "Carbon chain length",
       size = "Sub-family", color = "Sub-family")

ggplot_contam_profile_prop_PFAS

ggsave("../inst/results/Data_analysis/8.contam_profiles_prop_PFAS.jpg",
       ggplot_contam_profile_prop_PFAS, width = 6, height = 6, dpi = 300)
```

